### 2024-11-29 model_lba.Rmd

---
title: "model_lba"
author: "Sam & Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script sets up the LBA model, loads in the raw data, fits the model to the data and then samples the hLBA. There are tests at the bottom that compare parameter estimates.

Note - you need to make sure the working directory is the top directory with the models and dmc folder

# load data, models and packages #

```{r}
source ("dmc/dmc.R")
load_model ("LBA","lba_B.R") # LBA model with B=b-A parameterization

library(tidyverse)
```

Load in the model object if previously run

```{r}
## The model object is available upon request
## Please email Samantha Parker (samantha.parker@students.mq.edu.au) or Richard Ramsey (richard.ramsey@hest.ethz.ch)
# load("models/ea/modelKN.RData")
```

Load in the raw data (and convert names to match the labels that Sam used when fitting the LBA model)

Knots Data 
s: Subject ID (1-18)
S: Stimulus (match vs. mismatch)
GRP: Training Group (name vs. tie vs. both vs. untrained)
R: Response (same vs. different)
RT: Reaction Time (seconds)

For DMC, we also need to rename the levels of the response so that they are unique to stimulus.

```{r}
data_lba <- read_csv("data/processed/data_lba.csv") %>%
  ## rename group
  rename(GRP = GROUP) %>% 
  ## reorder
  select(s, S, GRP, R, RT) %>% 
  ## rename levels of variables
  mutate(S = if_else(S==1, "match", "mis"),
         R = if_else(R==1, "same", "diff")) %>% 
  ## create factors
  mutate(s = factor(s),
         S = factor(S,
                    levels = c("match", "mis")),
         GRP = factor(GRP,
                      levels = c("name", "tie", "both", "untrained")), ## this is the order Sam used.
         R = factor(R,
                    levels = c("same", "diff"))) %>% 
  as.data.frame() ## dmc prefers a data frame rather than a tibble
head(data_lba)
str(data_lba)
```

rename the data to match Sam's naming

```{r}
data_knots <- data_lba
```

# STEP 1: Set up the Model Object #

1) List the factors

In this design have a stimulus factor and a group factor (GRP)

```{r}
factors <- list(S=c("match","mis"), GRP=c("name","tie", "both", "untrained")) 
```

2) List of responses

```{r}
responses <- c("same","diff") # Can identify two pictures as the same or different
```

3) Match map
Specify which accumulator matches which response

```{r}
match.map <- list(M=list(match="same",mis="diff"))
```

4) p.map specifies how the model parameters match to the factors

In this design the threshold parameter (B) can be influenced by GRP and Response (R)
The drift rate parameter (mean_v) is influenced by GRP and Match
The non-decision time parameter (t0) can be influenced by the GRP factor
When specified as 1 this assumes that the parameter is the same for all accumulators

```{r}
p.map <- list(A="1",B=c("GRP", "R"),t0=c("GRP"),mean_v=c("GRP", "M"),sd_v="M",st0="1") #M needs to be put last
```

5) Constants

```{r}
const <- c(sd_v.false=1,st0=0)
```

6) Create data.model object

```{r}
model <- model.dmc(p.map,match.map=match.map,constants=const,responses=responses,
                   factors=factors)
```

Parameter vector names are: ( see attr(,"p.vector") )
 [1] "A"                      "B.name.same"            "B.tie.same"             "B.both.same"            "B.untrained.same"      
 [6] "B.name.diff"            "B.tie.diff"             "B.both.diff"            "B.untrained.diff"       "t0.name"               
[11] "t0.tie"                 "t0.both"                "t0.untrained"           "mean_v.name.true"       "mean_v.tie.true"       
[16] "mean_v.both.true"       "mean_v.untrained.true"  "mean_v.name.false"      "mean_v.tie.false"       "mean_v.both.false"     
[21] "mean_v.untrained.false" "sd_v.true"

Constants are (see attr(,"constants") ):
sd_v.false        st0 
         1          0

Model type = norm (posdrift= TRUE )

# STEP 2: COMBINE MODEL WITH DATA #

7) Simulate the data to make sure the model is set up correctly

```{r}
p.vector<- c(A = 1, 
             B.name.same = 0.5, 
             B.tie.same = 0.5, 
             B.both.same = 0.5, 
             B.untrained.same = 0.5, 
             B.name.diff = 0.5, 
             B.tie.diff = 0.5, 
             B.both.diff = 0.5, 
             B.untrained.diff = 0.5, 
             t0.name = .2, 
             t0.tie = .2, 
             t0.both = .2,
             t0.untrained = .2,
             mean_v.name.true = 1, 
             mean_v.tie.true = 1, 
             mean_v.both.true = 1, 
             mean_v.untrained.true = 1,
             mean_v.name.false = 0, 
             mean_v.tie.false = 0, 
             mean_v.both.false = 0, 
             mean_v.untrained.false = 0,
             sd_v.true = 0.5) #
check.p.vector(p.vector,model) #If everything ok nothing returned
print.cell.p(p.vector,model)

# Simulate this is for one subject
data <- simulate.dmc(p.vector,model,1e4)
data.model <- data.model.dmc(data,model)
plot.cell.density(data.model[data.model$S=='match',],C="same")
plot.cell.density(data.model[data.model$S=='match',],C="diff")

#This is for 20 subjects, 40 data points per subject
data1 <- h.simulate.dmc(model,ps=p.vector,ns=20,n=40)
data1.model <- data.model.dmc(data1,model)
```

8) Create the data model object

```{r}
data.model <- data.model.dmc(data_knots,model)
```

Plot the first two subjects to check

```{r}
 par(mfcol=c(2,4)) 
 for (i in 1:2) {
   plot.cell.density(data.cell=data.model[[i]][data.model[[i]]$GRP=="name" & data.model[[i]]$S=="match",],C="same")
   plot.cell.density(data.cell=data.model[[i]][data.model[[i]]$GRP=="name" & data.model[[i]]$S=="match",],C="diff")
   plot.cell.density(data.cell=data.model[[i]][data.model[[i]]$GRP=="tie" & data.model[[i]]$S=="mis",],C="mis")
   plot.cell.density(data.cell=data.model[[i]][data.model[[i]]$GRP=="tie" & data.model[[i]]$S=="mis",],C="same")
 }  
```

Save the data model object

```{r}
dmKN <- data.model
save(data.model, file = "data/ea/models/dmKN.RData")
```

# STEP 3: CREATE PRIORS #

Threshold parameters are 1
Start point parameters is 1
Rate parameters are 1 for matching accumulators and 0 for mismatching accumulators
sv for matching is .5
t0 is .2
t0 is truncated at 1 seconds, v is unbounded, A, B, and sd

```{r}
## Standard deviations of population priors should be broad. Means should be roughly plausible values. 
pop.mean <- c(A = 1, 
             B.name.same = 0.5, 
             B.tie.same = 0.5, 
             B.both.same = 0.5, 
             B.untrained.same = 0.5, 
             B.name.diff = 0.5, 
             B.tie.diff = 0.5, 
             B.both.diff = 0.5, 
             B.untrained.diff = 0.5, 
             t0.name = .2, 
             t0.tie = .2, 
             t0.both = .2,
             t0.untrained = .2,
             mean_v.name.true = 1, 
             mean_v.tie.true = 1, 
             mean_v.both.true = 1, 
             mean_v.untrained.true = 1,
             mean_v.name.false = 0, 
             mean_v.tie.false = 0, 
             mean_v.both.false = 0, 
             mean_v.untrained.false = 0,
             sd_v.true = 0.5) #
pop.scale <-c(A = 2, 
             B.name.same = 2, 
             B.tie.same = 2, 
             B.both.same = 2, 
             B.untrained.same = 2, 
             B.name.diff = 2, 
             B.tie.diff = 2, 
             B.both.diff = 2, 
             B.untrained.diff = 2, 
             t0.name = 2, 
             t0.tie = 2, 
             t0.both = 2,
             t0.untrained = 2,
             mean_v.name.true = 2, 
             mean_v.tie.true = 2, 
             mean_v.both.true = 2, 
             mean_v.untrained.true = 2,
             mean_v.name.false = 2, 
             mean_v.tie.false = 2, 
             mean_v.both.false = 2, 
             mean_v.untrained.false = 2,
             sd_v.true = 2) #
pop.prior <- prior.p.dmc(
  dists = rep("tnorm",22),
  p1=pop.mean,p2=pop.scale,
  lower=c(0,0,0,0,0,0,0,0,0,0.1,0.1,0.1,0.1,NA,NA,NA,NA,NA,NA,NA,NA,0),
  upper=c(NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,1)
)
```

Check the distributions

```{r}
par(mfcol=c(2,5)); for (i in names(pop.prior)) plot.prior(i,pop.prior)
```

# STEP 4: SAMPLING - FIT FIXED EFFECTS AND HIERARCHICAL #

1) Specify the priors

```{r}
p.prior <- prior.p.dmc(
  dists = rep("tnorm",22),
  p1=pop.mean,p2=pop.scale,
  lower=c(0,0,0,0,0,0,0,0,0,0.1,0.1,0.1,0.1,NA,NA,NA,NA,NA,NA,NA,NA,0),
  upper=c(NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,1)
)
```

2) Set up the sampling object called sKN

```{r}
sKN  <- h.samples.dmc(nmc = 120, p.prior, data.model)
```

## Automatic Script for sampling ##

Fits fixed effects first and uses these as starting appoints for the full hierarchical model

```{r}
tmp=load("models/ea/modelKN.RData")
cores=length(sKN)

# Fixed effects, comment out on subsequent runs
sKN <- h.RUN.dmc(sKN,cores=cores)
save(dmKN,sKN,file="models/ea/donesmodelKN.RData")
# Make hierarchical, comment out on subsequent runs
p.prior <- sKN[[1]]$p.prior
p1 <- get.p.vector(sKN[[1]])[names(p.prior)]
s.prior <- prior.p.dmc(p1=p1,p2=p1,
                       dists=rep("gamma",length(p1)))
pp.prior=list(p.prior,s.prior)
hstart <- make.hstart(sKN)
theta1 <- make.theta1(sKN)
hKN <- h.samples.dmc(nmc=100,p.prior,dmKN,pp.prior,
                    hstart.prior=hstart,theta1=theta1,thin=10)
rm(dmKN,sKN); gc()

# Hierarhcial fit, comment out all but the next line on subsequent runs
pp.prior <- attr(hKN,"hyper")$pp.prior
hKN  <- h.run.unstuck.dmc(hKN, p.migrate = .05, cores = cores)
save(hKN,file="models/ea/modelKN.RData")
hKN <- h.run.converge.dmc(h.samples.dmc(nmc=100, samples=hKN),
                      thorough=TRUE,nmc=50,cores=cores,finalrun=TRUE,finalI=250)
save(hKN,file="models/ea/modelKN.RData")

# Uncomment the following and make more copies to run more fits.
# hX <- h.run.dmc(h.samples.dmc(samples=hX,nmc=250),cores=cores,report=10)
# save(hX,file="models/ea/modelX.RData")

# Make diagnostics, tweak layout as desired
pdf("figures/ea/modelKN_chains.pdf",height=6,width = 8)
plot.dmc(hKN,hyper=TRUE,pll.chain=TRUE) 
plot.dmc(hKN,hyper=TRUE,layout=c(3,3))
plot.dmc(hKN,hyper=TRUE,layout=c(4,4),p.prior=pp.prior)
dev.off()

ppKN <- h.post.predict.dmc(hKN)
save(hKN,ppKN,file="models/ea/modelKN.RData")
pdf("figures/ea/KN_fit.pdf",height=6, width=8)
plot.pp.dmc(ppKN,model.legend = TRUE,layout=c(2,4))
dev.off()

h.IC.dmc(hKN,DIC=TRUE)

gelman.diag.dmc(hKN,hyper=TRUE)
gelman.diag.dmc(hKN)

parsKN <- summary.dmc(hKN,hyper=TRUE) # Summary of the hyper posterior sample
round(parsKN$quantiles[,c(1,3,5)],3)
parsKN <- summary.dmc(hKN) # Summary of the posterior samples for each participant.
save(hKN, ppKN, parsKN,file="models/ea/modelKN.RData")
```

# STEP 5: PARAMETER ESTIMATES #

List of parameters since you will use these names in the parameter tests

Parameter vector names are: ( see attr(,"p.vector") )
 [1] "A"                      "B.name.same"            "B.tie.same"             "B.both.same"            "B.untrained.same"      
 [6] "B.name.diff"            "B.tie.diff"             "B.both.diff"            "B.untrained.diff"       "t0.name"               
[11] "t0.tie"                 "t0.both"                "t0.untrained"           "mean_v.name.true"       "mean_v.tie.true"       
[16] "mean_v.both.true"       "mean_v.untrained.true"  "mean_v.name.false"      "mean_v.tie.false"       "mean_v.both.false"     
[21] "mean_v.untrained.false" "sd_v.true"

Constants are (see attr(,"constants") ):
sd_v.false        st0 
         1          0

## DRIFT RATES ##

Calculate subject average medians for drift rates for TRUE - FALSE drift rate

```{r}
fun <- function(x) {
  c(diff(x[c("mean_v.name.false","mean_v.name.true")]),
    diff(x[c("mean_v.tie.false","mean_v.tie.true")]),
    diff(x[c("mean_v.both.false", "mean_v.both.true")]),
    diff(x[c("mean_v.untrained.false", "mean_v.untrained.true")]))
}

pretty <-  c("name","tie", "both", "untrained") #Set the names

ci.dv <- subject.average.ci(hKN,fun=fun,pretty=pretty)
round(ci.dv,2)

#       name  tie both untrained
# 2.5%  2.01 2.13 2.74      1.85
# 50%   2.30 2.39 3.12      2.08
# 97.5% 2.60 2.67 3.53      2.31
```

Parameter test of difference

```{r}
#Compare training effects against untrained

#Name vs. untrained
fun <- function(x){
  name <- diff(x[c("mean_v.name.false","mean_v.name.true")])            
  untrained <- diff(x[c("mean_v.untrained.false","mean_v.untrained.true")])            
  diff <- (untrained-name)
  c(name,untrained,diff)
}

compare.p(hKN,show.plot=TRUE,pretty=c("name","untrained"),fun=fun,xlab="Average Rate")

#        name untrained contrast
# 2.5%  2.011     1.848   -0.596
# 50%   2.296     2.076   -0.222
# 97.5% 2.595     2.315    0.154
# p.gt.0 
#  0.126

#tie vs. untrained
fun <- function(x){
  tie <- diff(x[c("mean_v.tie.false","mean_v.tie.true")])            
  untrained <- diff(x[c("mean_v.untrained.false","mean_v.untrained.true")])            
  diff <- (untrained-tie)
  c(tie,untrained,diff)
}

compare.p(hKN,show.plot=TRUE,pretty=c("tie","untrained"),fun=fun,xlab="Average Rate")

#         tie untrained contrast
# 2.5%  2.129     1.848   -0.671
# 50%   2.390     2.076   -0.313
# 97.5% 2.666     2.315    0.043
# p.gt.0 
#  0.042

#both vs. untrained
fun <- function(x){
  both <- diff(x[c("mean_v.both.false","mean_v.both.true")])            
  untrained <- diff(x[c("mean_v.untrained.false","mean_v.untrained.true")])            
  diff <- (untrained-both)
  c(both,untrained,diff)
}

compare.p(hKN,show.plot=TRUE,pretty=c("both","untrained"),fun=fun,xlab="Average Rate")

#        both untrained contrast
# 2.5%  2.744     1.848   -1.511
# 50%   3.122     2.076   -1.044
# 97.5% 3.531     2.315   -0.606
# p.gt.0 
#      0

#both vs. name
fun <- function(x){
  both <- diff(x[c("mean_v.both.false","mean_v.both.true")])            
  name <- diff(x[c("mean_v.name.false","mean_v.name.true")])            
  diff <- (name-both)
  c(both,name,diff)
}

compare.p(hKN,show.plot=TRUE,pretty=c("both","name"),fun=fun,xlab="Average Rate")

#        both     name contrast
# 2.5%  2.744     2.011   -1.322
# 50%   3.122     2.296   -0.825
# 97.5% 3.531     2.595   -0.327
# p.gt.0 
#  0.001

#both vs. tie
fun <- function(x){
  both <- diff(x[c("mean_v.both.false","mean_v.both.true")])            
  tie <- diff(x[c("mean_v.tie.false","mean_v.tie.true")])            
  diff <- (tie-both)
  c(both,tie,diff)
}

compare.p(hKN,show.plot=TRUE,pretty=c("both","tie"),fun=fun,xlab="Average Rate")

#        both   tie contrast
# 2.5%  2.744 2.129   -1.204
# 50%   3.122 2.390   -0.734
# 97.5% 3.531 2.666   -0.262
# p.gt.0 
#  0.001

```

## THRESHOLDS ##

Thresholds

 "B.name.same"            "B.tie.same"             "B.both.same"            "B.untrained.same"      
 "B.name.diff"            "B.tie.diff"             "B.both.diff"            "B.untrained.diff"

Not interested in these differences but check anyway.
Look at B for same v diff collapsing over training type
```{r}
fun <- function(x){
  name.same <- (x[c( "B.name.same")])  
  tie.same <- (x[c("B.tie.same")])             
  both.same <- (x[c("B.both.same")])             
  untrained.same <- (x[c("B.untrained.same")])             
  name.diff <- (x[c("B.name.diff")])    
  tie.diff <- (x[c("B.tie.diff")]) 
  both.diff <- (x[c("B.both.diff")])  
  untrained.diff <- (x[c("B.untrained.diff")])             
  c(B.same <- (name.same+tie.same+both.same+untrained.same)/4,
  B.diff <- (name.diff+tie.diff+both.diff+untrained.diff)/4)
}

pretty <-  c("B.same","B.diff") #Set the names

ci.dv <- subject.average.ci(hKN,fun=fun,pretty=pretty)
round(ci.dv,2)

#       B.same B.diff
# 2.5%    1.03   0.78
# 50%     1.11   0.84
# 97.5%   1.19   0.90

#Parameter test to compare B for RJA vs. Con
fun <- function(x){
  name.same <- (x[c( "B.name.same")])  
  tie.same <- (x[c("B.tie.same")])             
  both.same <- (x[c("B.both.same")])             
  untrained.same <- (x[c("B.untrained.same")])             
  name.diff <- (x[c("B.name.diff")])    
  tie.diff <- (x[c("B.tie.diff")]) 
  both.diff <- (x[c("B.both.diff")])  
  untrained.diff <- (x[c("B.untrained.diff")])             
  B.same <- (name.same+tie.same+both.same+untrained.same)/4
  B.diff <- (name.diff+tie.diff+both.diff+untrained.diff)/4
  c(B.same, B.diff, B.same - B.diff)
}
compare.p(hKN,show.plot=TRUE,pretty=c("BSame","BDiff"),fun=fun,xlab="Threshold Same v Diff")

#       BSame BDiff contrast
# 2.5%  1.035 0.778    0.195
# 50%   1.110 0.840    0.270
# 97.5% 1.188 0.903    0.346
# p.gt.0 
#      1 
```

Check B collapsing across same and diff by training condition

```{r}
fun <- function(x){
  name.same <- (x[c( "B.name.same")])  
  tie.same <- (x[c("B.tie.same")])             
  both.same <- (x[c("B.both.same")])             
  untrained.same <- (x[c("B.untrained.same")])             
  name.diff <- (x[c("B.name.diff")])    
  tie.diff <- (x[c("B.tie.diff")]) 
  both.diff <- (x[c("B.both.diff")])  
  untrained.diff <- (x[c("B.untrained.diff")])             
  c(B.name <- (name.same+name.diff)/2,
  B.tie <- (tie.same+tie.diff)/2,
  B.both <- (both.same+both.diff/2),
  B.untrained <- (untrained.same+untrained.diff/2))
}

pretty <-  c("B.name","B.tie", "B.both", "B.untrained") #Set the names

ci.dv <- subject.average.ci(hKN,fun=fun,pretty=pretty)
round(ci.dv,2)

#       B.name B.tie B.both B.untrained
# 2.5%    0.64  0.57   0.84        2.60
# 50%     0.76  0.68   1.01        2.80
# 97.5%   0.88  0.79   1.19        3.03

#Parameter Tests

#Name vs. untrained
fun <- function(x){
  name.same <- (x[c( "B.name.same")])  
  tie.same <- (x[c("B.tie.same")])             
  both.same <- (x[c("B.both.same")])             
  untrained.same <- (x[c("B.untrained.same")])             
  name.diff <- (x[c("B.name.diff")])    
  tie.diff <- (x[c("B.tie.diff")]) 
  both.diff <- (x[c("B.both.diff")])  
  untrained.diff <- (x[c("B.untrained.diff")])             
  B.name <- (name.same+name.diff)/2
  B.untrained <- (untrained.same+untrained.diff)/2
  c(B.name, B.untrained, B.name - B.untrained)
}
compare.p(hKN,show.plot=TRUE,pretty=c("BName","BUntrained"),fun=fun,xlab="Threshold Name v Untrained")

#       BName BUntrained contrast
# 2.5%  0.635      1.702   -1.250
# 50%   0.756      1.829   -1.075
# 97.5% 0.880      1.970   -0.894
# p.gt.0 
#      0

#Tie vs. untrained
fun <- function(x){
  name.same <- (x[c( "B.name.same")])  
  tie.same <- (x[c("B.tie.same")])             
  both.same <- (x[c("B.both.same")])             
  untrained.same <- (x[c("B.untrained.same")])             
  name.diff <- (x[c("B.name.diff")])    
  tie.diff <- (x[c("B.tie.diff")]) 
  both.diff <- (x[c("B.both.diff")])  
  untrained.diff <- (x[c("B.untrained.diff")])             
  B.tie <- (tie.same+tie.diff)/2
  B.untrained <- (untrained.same+untrained.diff)/2
  c(B.tie, B.untrained, B.tie - B.untrained)
}
compare.p(hKN,show.plot=TRUE,pretty=c("BTie","BUntrained"),fun=fun,xlab="Threshold Tie v Untrained")

#        BTie BUntrained contrast
# 2.5%  0.573      1.702   -1.332
# 50%   0.677      1.829   -1.150
# 97.5% 0.787      1.970   -0.985
# p.gt.0 
#      0

#Both vs. untrained
fun <- function(x){
  name.same <- (x[c( "B.name.same")])  
  tie.same <- (x[c("B.tie.same")])             
  both.same <- (x[c("B.both.same")])             
  untrained.same <- (x[c("B.untrained.same")])             
  name.diff <- (x[c("B.name.diff")])    
  tie.diff <- (x[c("B.tie.diff")]) 
  both.diff <- (x[c("B.both.diff")])  
  untrained.diff <- (x[c("B.untrained.diff")])             
  B.both <- (both.same+both.diff/2)
  B.untrained <- (untrained.same+untrained.diff)/2
  c(B.both, B.untrained, B.both - B.untrained)
}
compare.p(hKN,show.plot=TRUE,pretty=c("BBoth","BUntrained"),fun=fun,xlab="Threshold Both v Untrained")

#       BBoth BUntrained contrast
# 2.5%  0.845      1.702   -1.041
# 50%   1.012      1.829   -0.815
# 97.5% 1.189      1.970   -0.595
# p.gt.0 
#      0

#Both vs. name
fun <- function(x){
  name.same <- (x[c( "B.name.same")])  
  tie.same <- (x[c("B.tie.same")])             
  both.same <- (x[c("B.both.same")])             
  untrained.same <- (x[c("B.untrained.same")])             
  name.diff <- (x[c("B.name.diff")])    
  tie.diff <- (x[c("B.tie.diff")]) 
  both.diff <- (x[c("B.both.diff")])  
  untrained.diff <- (x[c("B.untrained.diff")])             
  B.both <- (both.same+both.diff/2)
  B.name <- (name.same+name.diff)/2
  c(B.both, B.name, B.both - B.name)
}
compare.p(hKN,show.plot=TRUE,pretty=c("BBoth","BName"),fun=fun,xlab="Threshold Both v Name")

#      BBoth BName contrast
# 2.5%  0.845 0.635    0.042
# 50%   1.012 0.756    0.256
# 97.5% 1.189 0.880    0.477
# p.gt.0 
#   0.99

#Both vs. tie
fun <- function(x){
  name.same <- (x[c( "B.name.same")])  
  tie.same <- (x[c("B.tie.same")])             
  both.same <- (x[c("B.both.same")])             
  untrained.same <- (x[c("B.untrained.same")])             
  name.diff <- (x[c("B.name.diff")])    
  tie.diff <- (x[c("B.tie.diff")]) 
  both.diff <- (x[c("B.both.diff")])  
  untrained.diff <- (x[c("B.untrained.diff")])             
  B.both <- (both.same+both.diff/2)
  B.tie <- (tie.same+tie.diff)/2
  c(B.both, B.tie, B.both - B.tie)
}
compare.p(hKN,show.plot=TRUE,pretty=c("BBoth","BTie"),fun=fun,xlab="Threshold Both v Tie")

#       BBoth  BTie contrast
# 2.5%  0.845 0.573    0.136
# 50%   1.012 0.677    0.335
# 97.5% 1.189 0.787    0.538
# p.gt.0 
#      1

#Name v Tie
fun <- function(x){
  name.same <- (x[c( "B.name.same")])  
  tie.same <- (x[c("B.tie.same")])             
  both.same <- (x[c("B.both.same")])             
  untrained.same <- (x[c("B.untrained.same")])             
  name.diff <- (x[c("B.name.diff")])    
  tie.diff <- (x[c("B.tie.diff")]) 
  both.diff <- (x[c("B.both.diff")])  
  untrained.diff <- (x[c("B.untrained.diff")])             
  B.name <- (name.same+name.diff)/2
  B.tie <- (tie.same+tie.diff)/2
  c(B.name, B.tie, B.name - B.tie)
}
compare.p(hKN,show.plot=TRUE,pretty=c("Bname","BTie"),fun=fun,xlab="Threshold Name v Tie")

#       Bname  BTie contrast
# 2.5%  0.635 0.573   -0.087
# 50%   0.756 0.677    0.079
# 97.5% 0.880 0.787    0.240
# p.gt.0 
#  0.819

```

Parameter test of difference
"t0.name"               "t0.tie"                 "t0.both"                "t0.untrained"

```{r}
#Compare validity effect at each level of congruency

fun <- function(x){
  name <- (x[c( "t0.name")])  
  tie <- (x[c("t0.tie")])             
  both <- (x[c("t0.both")])             
  untrained <- (x[c("t0.untrained")])             
  c(t0.name <- (name),
  t0.tie <- (tie),
  t0.both <- (both),
  t0.untrained <- (untrained))
}

pretty <-  c("name","tie", "both", "untrained") #Set the names

ci.dv <- subject.average.ci(hKN,fun=fun,pretty=pretty)
round(ci.dv,2)

#       name  tie both untrained
# 2.5%  0.47 0.48 0.50      0.22
# 50%   0.50 0.51 0.52      0.26
# 97.5% 0.53 0.54 0.55      0.29

#Parameter tests 
# Name vs. untrained
compare.p(hKN,show.plot=TRUE,pnames=c("t0.name","t0.untrained"),pretty=c("name","untrained"),xlab="Average t0")

#        name untrained contrast
# 2.5%  0.469     0.222    0.194
# 50%   0.501     0.258    0.243
# 97.5% 0.533     0.295    0.291
# p.gt.0 
#      1

#Tie vs untrained
compare.p(hKN,show.plot=TRUE,pnames=c("t0.tie","t0.untrained"),pretty=c("tie","untrained"),xlab="Average t0")
#         tie untrained contrast
# 2.5%  0.481     0.222    0.205
# 50%   0.509     0.258    0.250
# 97.5% 0.536     0.295    0.298
# p.gt.0 
#      1

#Both vs. untrained
compare.p(hKN,show.plot=TRUE,pnames=c("t0.both","t0.untrained"),pretty=c("both","untrained"),xlab="Average t0")

#        both untrained contrast
# 2.5%  0.495     0.222    0.219
# 50%   0.524     0.258    0.266
# 97.5% 0.553     0.295    0.314
# p.gt.0 
#      1

```