### 2024-11-29 wrangle.Rmd

---
title: "wrangle"
author: "Sam & Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the main data wrangling file for the analysis of knots data from Cross et al. (2012).

This script loads the processed data, wrangles the data, makes summary plots and writes out dataframes that are appropriate for the different types of analyses that will follow in the modelling scripts.

# load libraries #

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork")

lapply(pkg, library, character.only = TRUE)
```

# plot settings #

theme settings for ggplot

```{r, eval = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18, face = "bold"), 
          title = element_text(size = 18, face = "bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# 1. load the data #

this is the processed data for day 5 session 2

```{r}
raw <- read_csv("data/processed/day5_post_knots_data.csv")
head(raw)
```

check only day 5, session 2

```{r}
raw %>%
  distinct(DAY, SESSION)
```

# 2. do some wrangling #

select relevant variables.
create an accuracy variable.
rename group.
create factors.

```{r}
data <- raw %>% 
  select(PID, GROUP, MATCH, RESPONSE, RT) %>% 
  mutate(ACC = if_else(MATCH == RESPONSE, 1, 0)) %>% 
  mutate(GROUP = case_when(
    GROUP == 1 ~ "name",
    GROUP == 2 ~ "tie",
    GROUP == 3 ~ "both",
    .default = "untrained")) %>% 
  mutate(PID = factor(PID),
         GROUP = factor(GROUP,
                        levels = c("untrained", "name", "tie", "both")))
head(data)
glimpse(data)
```

```{r}
# Relevant variables as follows:
# PID = ID each participants should be n = 18 (with day 5, session 2 data)
# RESPONSE = Whether the participant selects a match or not? (1 = match, 2 = mismatch)
# RT = Response times in ms
# GROUP = 1-4 (name, tie, both, untrained)
# MATCH = Whether they match or not (1 = match and 2 = mismatch)
# ACC = correct = 1, incorrect = 0
```

create a reaction time dataframe

```{r}
data_rt <- data %>% 
  filter(ACC==1)
data_rt
```

# 3. take a quick look at the data #

trial count per PID

```{r}
trials_pid <- data %>%
  group_by(PID) %>%
  tally() %>% 
  arrange(n) # take a look at the lowest trial count
head(trials_pid)
```

per condition

```{r}
trials_cond <- data %>%
  group_by(GROUP) %>%
  tally() %>% 
  arrange(n) # take a look at the lowest trial count
head(trials_cond)
```

per PID and condition

```{r}
trials_pid_cond <- data %>%
  group_by(PID, GROUP) %>%
  tally() %>% 
  arrange(n) 
head(trials_pid_cond)
```

and print it

```{r}
trials_pid_cond %>% 
  print(n=Inf)
```

# 4. create summary data #

## accuracy ##

at the pid level

```{r}
data_s_pid <- data %>% 
  group_by(PID, GROUP) %>% 
  summarise(sum_acc = sum(ACC), 
            mean_acc = mean(ACC, na.rm = TRUE) * 100, 
            sd_acc = sd(ACC, na.rm = TRUE) * 100)
head(data_s_pid)
```

at the group level

```{r}
data_s_g <- data %>% 
  group_by(GROUP) %>% 
  summarise(n=length(unique(PID)),
            sum_acc = sum(ACC), 
            mean_acc = mean(ACC, na.rm = TRUE) * 100, 
            sd_acc = sd(ACC, na.rm = TRUE) * 100, 
            sem_acc = (sd_acc/sqrt(n)))
data_s_g
```

## reaction time ##

at the pid level

```{r}
data_s_rt_pid <- data_rt %>% 
  group_by(PID, GROUP) %>%
  summarise(mean_rt = mean(RT, na.rm = TRUE), 
            sd_rt = sd(RT, na.rm = TRUE))
head(data_s_rt_pid)
```

at the group level

```{r}
data_s_rt_g <- data_rt %>% 
  group_by(GROUP) %>%
  summarise(n=length(unique(PID)),
            mean_rt = mean(RT, na.rm = TRUE), 
            sd_rt = sd(RT, na.rm = TRUE), 
            sem_rt = (sd_rt/sqrt(n)))
data_s_rt_g
```

# 5. make some plots #

plot accuracy

```{r}
p5.1 <- ggplot(data_s_pid, aes(x = GROUP, y = mean_acc, 
                               fill = GROUP, colour = GROUP)) +
  geom_jitter(position=position_jitterdodge(dodge.width =1), 
              size = 2.5, colour="grey") +
  geom_violin(alpha = 0.5, position=pd2) +
  geom_errorbar(data = data_s_g, aes(ymin=mean_acc-sem_acc, ymax=mean_acc+sem_acc), 
                width=.3, size =0.5, colour = "black", position=pd2) +
  geom_point(data = data_s_g, colour = "black", position=pd2, size =4) +
  geom_line(data=data_s_g, group=1, colour = "black") +
  xlab("Training Condition") + 
  ylab('Percentage Correct') + 
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  theme(legend.position = "none")
p5.1
```

plot rt

```{r}
p5.2 <- ggplot(data_s_rt_pid, aes(x = GROUP, y = mean_rt, 
                               fill = GROUP, colour = GROUP)) +
  geom_jitter(position=position_jitterdodge(dodge.width =1), 
              size = 2.5, colour="grey") +
  geom_violin(alpha = 0.5, position=pd2) +
  geom_errorbar(data = data_s_rt_g, aes(ymin=mean_rt-sem_rt,
                                        ymax=mean_rt+sem_rt), 
                width=.3, size =0.5, colour = "black", position=pd2) +
  geom_point(data = data_s_rt_g, colour = "black", position=pd2, size =4) +
  geom_line(data=data_s_rt_g, group=1, colour = "black") +
  xlab("Training Condition") + 
  ylab('Mean Reaction Time (ms)') + 
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  theme(legend.position = "none")
p5.2
```

plot together

```{r}
p5.3 <- (p5.1 | p5.2) +
  plot_annotation(tag_levels = 'A')
p5.3

ggsave("figures/descriptive/violin_acc_rt.jpeg",
       width = 10, height = 6, dpi = 800)
```

# 6. write out the data #

## data for plotting and modelling via brms ##

this is the data for acc and rt labelled as it is in plots
It can be used for modelling in the manifest analysis of rt and acc separately

```{r}
write_csv(data, "data/processed/data_acc.csv")
write_csv(data_rt, "data/processed/data_rt.csv")
```

## data for use in LBA modelling ##

this is for modelling with DMC software when using an LBA model.
DMC has specific rules about how the data are structured.

Rules for data to be used in dmc

1) Make sure you have an "s" column identifying the data for each subject that is a factor
2) Follow that by columns specifying the factors in your design. Again each must be a factor
3) The second last column give the response (again a factor)
4) The last columns is called RT, make sure the unit is seconds.

And we can remove the ACC column as DMC makes the match between Stimulus and Response.

```{r}
data_lba <- data %>%
  rename(s = PID, S = MATCH, R = RESPONSE) %>% 
  mutate(RT = RT/1000,
         S = factor(S),
         R = factor(R)) %>%
  select(s, GROUP, S, R, RT) %>% 
  as.data.frame()
head(data_lba)

## technically, since we are saving out a .csv file then reading it in when we get to the model script the factors and data frame (vs tibble) are not necessary here as they will need to be re-done when they are read in later. But I think it is probably good practice to get this in the format that is later required.
```

save it as a .csv file

```{r}
write_csv(data_lba, "data/processed/data_lba.csv")
```