### 2024-12-06 effects_rt_acc.Rmd
added a few new plots and tables for supps

---
title: "effects_rt_acc"
author: "Sam & Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

Secondary analysis of manifest measures in knots data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file visualises and tabulates parameters in the most complex model from the manifest analysis of the knots data.

This file focusses on the effects from reaction time and accuracy models.

# load libraries #

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "brms", "tidybayes", 
         "ggdist", "future", "bayesplot", "parallel", "here", "tinytable",
         "webshot2", "rstan")

lapply(pkg, library, character.only = TRUE)
```

# plot settings #

theme settings for ggplot

```{r, eval = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18, face = "bold"), 
          title = element_text(size = 18, face = "bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# table options #

```{r}
options(tinytable_tt_digits = 2,
        tinytable_print_output = "markdown")
```

# 1. look at reaction time #

## 1.1 read in model object, fits and data (as necessary) ##

```{r}
# rt
b1 <- readRDS("models/manifest/b1.rds")
```

read in the data

```{r}
# rt data
data_rt <- read_csv("data/processed/data_rt.csv") %>% 
  mutate(PID=factor(PID),
         GROUP = factor(GROUP,
                        levels = c("untrained", "name", "tie", "both")))
head(data_rt)
```

## 1.2 Look at the point and interval estimates for the full models ##

### summarise point and interval estimates for all aspects of the model ###

using posterior_samples and tidybayes

```{r}
post <- as_draws_df(b1) %>% #
  select(-lp__) %>% 
  as_tibble()

post_summary <- posterior_summary(b1, robust = TRUE)

# all fixed effects
post_qi_b <- post %>%
  select(starts_with("b_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_b)

# average varying effects - SD
post_qi_sd <- post %>%
  select(starts_with("sd_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_sd)

# varying effects by participant 
# I take this to be 18 pts x 5 terms (ndt_intercept, intercept, 3 levels of group) = 90 in total
vary_vals <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(vary_vals)
```

### visualise fixed effects ###

Using posterior_samples and tidybayes

```{r}
## fixed effects of interest, excluding basic intercepts 
## Is this the right way to label the interaction?
tidy_fixed <- post %>% 
  select(starts_with("b_")) %>% # select and rename in simpler labels.
  pivot_longer(everything()) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(name, "ndt"), "ndt_intercept",
                if_else(str_detect(name, "name"), "name",
                if_else(str_detect(name, "tie"), "tie",
                if_else(str_detect(name, "both"), "both", "intercept")))),
         term = factor(term, levels=unique(term)))
head(tidy_fixed)
tail(tidy_fixed)

## check labels
check.labels <- tidy_fixed %>% 
  group_by(key, term) %>% 
  distinct(key, term)
check.labels

# remove intercepts as they are on a different scale and make it difficult to plot
tidy_fixed <- tidy_fixed %>%
  filter(!term %in% c("intercept", "ndt_intercept"))
head(tidy_fixed)

check.labels <- tidy_fixed %>% 
  group_by(key, term) %>% 
  distinct(key, term)
check.labels

p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = fct_rev(term), 
                                       fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(alpha=0.7) +
  labs(#title = "fixed effects of condition",
       x = "Log(RT)", y = NULL) +
  scale_fill_brewer(palette = "Dark2")
p_tidy_fixed

# ggsave ("figures/manifest/b1_fixef.jpeg",
#         width = 8, height = 6, dpi = 800)
```

### make a table of fixed effects ###

```{r}
b1_fixef_tab <- tidy_fixed %>%
  select(term, value) %>% 
  group_by(term) %>% 
  median_qi(value) %>% 
  mutate(dv = "reaction time") %>% 
  select(dv, term, value, .lower, .upper) 
b1_fixef_tab
```

### visualise varying effects ###

using posterior samples and tidybayes.

variance at the group level (i.e., intercepts and effects across participants)

```{r}
## random effects of interest, including item, as well as participant and effects across pts
tidy_varying <- post %>% 
  select(contains("sd_")) %>% 
  pivot_longer(everything()) %>% 
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(name, "ndt"), "ndt_intercept",
                if_else(str_detect(name, "name"), "name",
                if_else(str_detect(name, "tie"), "tie",
                if_else(str_detect(name, "both"), "both", "intercept")))),
         term = factor(term, levels=c("ndt_intercept", "intercept",
                                      "name", "tie", "both")))
head(tidy_varying)
tail(tidy_varying)
str(tidy_varying)

check.varying <- tidy_varying %>% 
  group_by(key, term) %>% 
  distinct(key, term)
check.varying

p_tidy_varying_pid <- ggplot(tidy_varying, aes(x = value, y = fct_rev(term))) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(alpha = 0.7) +
  labs(title = "SD plot for varying participant effects",
       x = NULL, y = NULL) +
  coord_cartesian(xlim =c(0, 0.5))+
  scale_x_continuous(breaks=seq(0,1,0.1))
p_tidy_varying_pid

ggsave ("figures/manifest/b1_varying_pid.jpeg",
        width = 8, height = 6, dpi = 800)
```

### visualise variance within group levels ###

i.e., at the individual participant level

Note - this should be in a centred metric so that deviation from the group average (zero) can be visualised. All good, but interpreting intercepts in some cases can difficult because it is when all other factors are zero. It can therefore be a little difficult to interpret and they may not be completely symmetric around zero for that reason.

use output from posterior summary and wrangle

```{r}
# first filter and create new tibbles
# mutate here to create factors to preserve order of term - useful when plotting. Also create a new outlier variable and factor it. You could do the outlier variable in one step possibly using as_factor, but this does the job.

# participant level
vary_pt <- post_summary %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_PID"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(term = if_else(str_detect(param, "ndt"), "ndt_intercept",
                if_else(str_detect(param, "name"), "name",
                if_else(str_detect(param, "tie"), "tie",
                if_else(str_detect(param, "both"), "both", "intercept")))),
        term = factor(term, levels=c("ndt_intercept", "intercept",
                                      "name", "tie", "both"))) %>%  
  arrange(term, Estimate) %>% 
  mutate(rank = rep(1:18, times=5),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
         outlier = factor(outlier, levels = c("no", "yes")))
head(vary_pt)
```

plot varying effects via caterpillar plots

variation by participant

```{r}
p_tidy_varyvals <- ggplot(vary_pt, aes(x=rank, y=Estimate,
                                           colour=outlier))+
  geom_pointrange(aes(ymin=Q2.5,ymax=Q97.5),fatten = 0.5)+
  geom_hline(yintercept=0, colour = "red")+
  scale_colour_manual(values=c("darkgrey","black")) +
  labs(title = "Varying effects by participant",
       x = "participant id")+
  facet_wrap(~term) +
  theme(legend.position = "none")
p_tidy_varyvals

ggsave ("figures/manifest/b1_varyvals.jpeg",
        width = 8, height = 6, dpi = 800)
```

## 1.3 plot in the original metric as a function of conditions ##

This can be useful to see what the model predicts in terms of reaction time or points on a scale, rather than parameter estimates. This is especially relevant for more complex models like shifted log normal models where individual parameters may be hard to interpret on their own.

### posterior predictions ###

calculate

```{r}
ppreds_rt <- data_rt %>%
  distinct(PID, GROUP) %>%
  add_epred_draws(b1) 
head(ppreds_rt)
tail(ppreds_rt)
```

summary

```{r}
ppreds_rt_g <- ppreds_rt %>% 
  group_by(GROUP, .draw) %>% 
  summarise(.epred = mean(.epred))
ppreds_rt_g
```

plot

```{r}
p_ppreds_rt <- ggplot(ppreds_rt_g, aes(x = GROUP, y = .epred, 
                                       fill=GROUP)) +  
  stat_halfeye(alpha=0.7) +
  labs(x = "condition", y = "RT (ms)") +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "none")
p_ppreds_rt
```

calculate quantiles

```{r}
ppreds_rt_q <- ppreds_rt_g %>% 
  group_by(GROUP) %>% 
  median_qi(.epred)
ppreds_rt_q
```

make a table

```{r}
ppred_rt_tab <- ppreds_rt_q %>%
  select(GROUP, .epred, .lower, .upper) %>% 
  rename(condition = GROUP)
ppred_rt_tab

# ## make a table using tinytable
tt_rt_tab <- tt(ppred_rt_tab,
                caption="Posterior predictions for RT") %>% 
  theme_tt("striped")
tt_rt_tab

## save it as a png
save_tt(tt_rt_tab, here("tables/manifest", "b1_ppreds_rt.png"),
        overwrite = T)
```

# 2. look at accuracy #

## 2.1 read in model object, fits and data (as necessary) ##

```{r}
# acc
ab1 <- readRDS("models/manifest/ab1.rds")
```

read in the data

```{r}
# acc data
data_acc <- read_csv("data/processed/data_acc.csv") %>% 
    mutate(PID=factor(PID),
           GROUP = factor(GROUP,
                          levels = c("untrained", "name", "tie", "both")))
head(data_acc)
```

## 2.2 Look at the point and interval estimates for the full models ##

### summarise point and interval estimates for all aspects of the model ###

using posterior_samples and tidybayes

```{r}
ab1_post <- as_draws_df(ab1) %>% #
  select(-lp__) %>% 
  as_tibble()

ab1_post_summary <- posterior_summary(ab1, robust = TRUE)

# all fixed effects
ab1_post_qi_b <- ab1_post %>%
  select(starts_with("b_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(ab1_post_qi_b)

# average varying effects - SD
ab1_post_qi_sd <- ab1_post %>%
  select(starts_with("sd_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(ab1_post_qi_sd)

# varying effects by participant 
# I take this to be 18 pts x 4 terms (intercept, 3 levels of group) = 72 in total
ab1_vary_vals <- ab1_post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(ab1_vary_vals)
```

### visualise fixed effects ###

Using posterior_samples and tidybayes

```{r}
## fixed effects of interest, excluding basic intercepts 
## Is this the right way to label the interaction?
ab1_tidy_fixed <- ab1_post %>% 
  select(starts_with("b_")) %>% # select and rename in simpler labels.
  pivot_longer(everything()) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(name, "name"), "name",
                if_else(str_detect(name, "tie"), "tie",
                if_else(str_detect(name, "both"), "both", "intercept"))),
         term = factor(term, levels=unique(term)))
head(ab1_tidy_fixed)
tail(ab1_tidy_fixed)

## check labels
check.labels <- ab1_tidy_fixed %>% 
  group_by(key, term) %>% 
  distinct(key, term)
check.labels

# remove intercepts as they are on a different scale and make it difficult to plot
ab1_tidy_fixed <- ab1_tidy_fixed %>%
  filter(!term %in% c("intercept"))
head(ab1_tidy_fixed)

check.labels <- ab1_tidy_fixed %>% 
  group_by(key, term) %>% 
  distinct(key, term)
check.labels

p_tidy_fixed_ab1 <- ggplot(ab1_tidy_fixed, aes(x = value, y = fct_rev(term), 
                                       fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(alpha=0.7) +
  labs(#title = "fixed effects of condition",
       x = "Accuracy", y = NULL) +
  scale_fill_brewer(palette = "Dark2")
p_tidy_fixed_ab1

# ggsave ("figures/manifest/ab1_fixef.jpeg",
#         width = 8, height = 6, dpi = 800)
```

### make a table of fixed effects ###

```{r}
ab1_fixef_tab <- ab1_tidy_fixed %>%
  select(term, value) %>% 
  group_by(term) %>% 
  median_qi(value) %>% 
  mutate(dv = "accuracy") %>% 
  select(dv, term, value, .lower, .upper) 
ab1_fixef_tab
```

### visualise varying effects ###

using posterior samples and tidybayes.

variance at the group level (i.e., intercepts and effects across participants)

```{r}
## random effects of interest, including item, as well as participant and effects across pts
ab1_tidy_varying <- ab1_post %>% 
  select(contains("sd_")) %>% 
  pivot_longer(everything()) %>% 
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(name, "name"), "name",
                if_else(str_detect(name, "tie"), "tie",
                if_else(str_detect(name, "both"), "both", "intercept"))),
         term = factor(term, levels=c("intercept",
                                      "name", "tie", "both")))
head(ab1_tidy_varying)
tail(ab1_tidy_varying)
str(ab1_tidy_varying)

check.varying <- ab1_tidy_varying %>% 
  group_by(key, term) %>% 
  distinct(key, term)
check.varying

p_tidy_varying_pid_ab1 <- ggplot(ab1_tidy_varying, aes(x = value, 
                                               y = fct_rev(term))) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(alpha = 0.7) +
  labs(title = "SD plot for varying participant effects",
       x = NULL, y = NULL) +
  coord_cartesian(xlim =c(0, 1))+
  scale_x_continuous(breaks=seq(0,1,0.1))
p_tidy_varying_pid_ab1

ggsave ("figures/manifest/ab1_varying_pid.jpeg",
        width = 8, height = 6, dpi = 800)
```

### visualise variance within group levels ###

i.e., at the individual participant level

Note - this should be in a centred metric so that deviation from the group average (zero) can be visualised. All good, but interpreting intercepts in some cases can difficult because it is when all other factors are zero. It can therefore be a little difficult to interpret and they may not be completely symmetric around zero for that reason.

use output from posterior summary and wrangle

```{r}
# first filter and create new tibbles
# mutate here to create factors to preserve order of term - useful when plotting. Also create a new outlier variable and factor it. You could do the outlier variable in one step possibly using as_factor, but this does the job.

# participant level
ab1_vary_pt <- ab1_post_summary %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_PID"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(term = if_else(str_detect(param, "name"), "name",
                if_else(str_detect(param, "tie"), "tie",
                if_else(str_detect(param, "both"), "both", "intercept"))),
         term = factor(term, levels=c("intercept",
                                      "name", "tie", "both"))) %>%  
  arrange(term, Estimate) %>% 
  mutate(rank = rep(1:18, times=4),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
         outlier = factor(outlier, levels = c("no", "yes")))
head(ab1_vary_pt)
```

plot varying effects via caterpillar plots

variation by participant

```{r}
p_tidy_varyvals_ab1 <- ggplot(ab1_vary_pt, aes(x=rank, y=Estimate,
                                           colour=outlier))+
  geom_pointrange(aes(ymin=Q2.5,ymax=Q97.5),fatten = 0.5)+
  geom_hline(yintercept=0, colour = "red")+
  scale_colour_manual(values=c("darkgrey","black")) +
  labs(title = "Varying effects by participant",
       x = "participant id")+
  facet_wrap(~term) +
  theme(legend.position = "none")
p_tidy_varyvals_ab1

ggsave ("figures/manifest/ab1_varyvals.jpeg",
        width = 8, height = 6, dpi = 800)
```

## 2.3 plot in the original metric as a function of conditions ##

This can be useful to see what the model predicts in terms of reaction time or points on a scale, rather than parameter estimates. This is especially relevant for more complex models like shifted log normal models where individual parameters may be hard to interpret on their own.

### posterior predictions ###

calculate

```{r}
ppreds_acc <- data_acc %>%
  distinct(PID, GROUP) %>%
  add_epred_draws(ab1) 
head(ppreds_acc)
tail(ppreds_acc)
```

summary

```{r}
ppreds_acc_g <- ppreds_acc %>% 
  group_by(GROUP, .draw) %>% 
  summarise(.epred = mean(.epred))
ppreds_acc_g
```

plot

```{r}
p_ppreds_acc <- ggplot(ppreds_acc_g, aes(x = GROUP, y = .epred, 
                                       fill=GROUP)) +  
  stat_halfeye(alpha=0.7) +
  labs(x = "condition", y = "accuracy (%)") +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "none")
p_ppreds_acc
```

calculate quantiles

```{r}
ppreds_acc_q <- ppreds_acc_g %>% 
  group_by(GROUP) %>% 
  median_qi(.epred)
ppreds_acc_q
```

make a table

```{r}
ppred_acc_tab <- ppreds_acc_q %>%
  select(GROUP, .epred, .lower, .upper) %>% 
  rename(condition = GROUP)
ppred_acc_tab

# ## make a table using tinytable
tt_acc_tab <- tt(ppred_acc_tab,
                caption="Posterior predictions for accuracy") %>% 
  theme_tt("striped")
tt_acc_tab

## save it as a png
save_tt(tt_acc_tab, here("tables/manifest", "ab1_ppreds_acc.png"),
        overwrite = T)
```

# 3. make plots for RT and Accuracy together #

fixed effects

```{r}
p_fixef <- (p_tidy_fixed_ab1 | p_tidy_fixed) +
  plot_annotation(title = "Fixed effects of condition for accuracy and RT",
                  tag_levels = 'A') +
  plot_layout(guides = 'collect',
              axes = 'collect')
p_fixef

ggsave("figures/manifest/fixef.jpeg",
       width = 10, height = 6, dpi = 800)
```

posterior predictions

```{r}
p_ppreds <- (p_ppreds_acc | p_ppreds_rt) +
  plot_annotation(title = "Posterior predictions for accuracy and RT",
                  tag_levels = 'A')
p_ppreds

ggsave("figures/manifest/ppreds.jpeg",
       width = 10, height = 6, dpi = 800)
```

# 4. make a table of fixed effects for rt and accuracy #

```{r}
fix_tab <- bind_rows(ab1_fixef_tab, b1_fixef_tab) %>% 
   mutate(across(where(is.double), \(x) round(x, 2)))
fix_tab

write_csv(fix_tab, "tables/manifest/fixef.csv")
```