### 2025-04-29 sim_data.Rmd
added a script to simulate data for code checking purposes

---
title: "sim_data"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file takes fitted regression models from the manifest analysis and uses them to simulate RT and accuracy data that has similar properties to the real data.

Since we cannot share the original data publicly because the ethical approval was not in place to do so when the data were collected in 2006, the purpose of this simulated data is to allow others to check the coding pipeline that we used.

Therefore, using this data will NOT produce the same results as we report in the manuscript, since the data are different. But the simulated data should resemble the structure of the real data.

Also note that for simplicity, we just use the manifest regression models and not the LBA models to generate reaction time and accuracy data. The reason for doing so is that our main aim is to provide fairly similar data to the original so that folks can reproduce our analytical pipeline, rather than any other aim that might rely on the specifics of using a LBA model.

# load libraries #

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "brms", "tidybayes", 
         "ggdist", "future", "bayesplot", "cmdstanr", "parallel", "rstan")

lapply(pkg, library, character.only = TRUE)
```

# plot settings #

theme settings for ggplot

```{r, eval = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18, face = "bold"), 
          title = element_text(size = 18, face = "bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# package settings #

```{r}
options(brms.backend = "cmdstanr",
        mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

rstan_options(auto_write = TRUE)

supportsMulticore()

detectCores()
```

# 1. load the models and the raw (original) data #

models

```{r}
## rt
b1 <- readRDS("models/manifest/b1.rds")

## accuracy
ab1 <- readRDS("models/manifest/ab1.rds")
```

data

```{r}
# rt data
data_rt <- read_csv("data/processed/data_rt.csv") %>% 
  mutate(PID=factor(PID),
         GROUP = factor(GROUP,
                        levels = c("untrained", "name", "tie", "both")))
head(data_rt)

# accuracy data
data_acc <- read_csv("data/processed/data_acc.csv") %>% 
    mutate(PID=factor(PID),
           GROUP = factor(GROUP,
                          levels = c("untrained", "name", "tie", "both")))
head(data_acc)
```

# 2. simulate accuracy data #

using the original data to define the design to make predictions on.

```{r}
# Step 1: Use the original data as the base for simulation
acc_preds <- data_acc %>%
  # This adds predicted values based on the posterior distribution
  add_predicted_draws(ab1,
                      seed = 123)
head(acc_preds)
glimpse(acc_preds)

# Step 2: For a single complete simulated dataset with same structure as original
acc_sim_data <- acc_preds %>%
  group_by(.row) %>%
  slice_sample(n = 1) %>%  # Take one random prediction per original observation
  ungroup()
head(acc_sim_data)
glimpse(acc_sim_data)
```

## create summary data ##

at the pid level

```{r}
acc_sim_pid <- acc_sim_data %>% 
  group_by(PID, GROUP) %>% 
  summarise(sum_acc = sum(.prediction), 
            mean_acc = mean(.prediction, na.rm = TRUE) * 100, 
            sd_acc = sd(.prediction, na.rm = TRUE) * 100)
head(acc_sim_pid)
```

at the group level

```{r}
acc_sim_g <- acc_sim_data %>% 
  group_by(GROUP) %>% 
  summarise(n=length(unique(PID)),
            sum_acc = sum(.prediction), 
            mean_acc = mean(.prediction, na.rm = TRUE) * 100, 
            sd_acc = sd(.prediction, na.rm = TRUE) * 100, 
            sem_acc = (sd_acc/sqrt(n)))
acc_sim_g

# # A tibble: 4 × 6
#   GROUP         n sum_acc mean_acc sd_acc sem_acc
#   <fct>     <int>   <int>    <dbl>  <dbl>   <dbl>
# 1 untrained    18     298     85.6   35.1    8.28
# 2 name         18     290     82.6   37.9    8.94
# 3 tie          18     312     89.1   31.2    7.34
# 4 both         18     308     87.7   32.8    7.74
```

## plot accuracy ##

```{r}
p2.1 <- ggplot(acc_sim_pid, aes(x = GROUP, y = mean_acc, 
                               fill = GROUP, colour = GROUP)) +
  geom_jitter(position=position_jitterdodge(dodge.width =1), 
              size = 2.5, colour="grey") +
  geom_violin(alpha = 0.5, position=pd2) +
  geom_errorbar(data = acc_sim_g, aes(ymin=mean_acc-sem_acc,
                                         ymax=mean_acc+sem_acc), 
                width=.3, linewidth =0.5, colour = "black", position=pd2) +
  geom_point(data = acc_sim_g, colour = "black", position=pd2, size =4) +
  geom_line(data=acc_sim_g, group=1, colour = "black") +
  xlab("Training Condition") + 
  ylab('Percentage Correct') + 
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  theme(legend.position = "none")
p2.1
```

That looks reasonably similar to the descriptive violin plots in the main text.

# 3. simulate reaction time data #

using the original data

```{r}
# Step 1: Use the original data as the base for simulation
# note - we still use the full data here data_acc, to make sure we get predictions for the full dataset, rather than just correct trials. But now we use the RT model to make predictions in terms of RT rather than accuracy.

rt_preds <- data_acc %>%
  # This adds predicted values based on the posterior distribution
  add_predicted_draws(b1,
                      seed = 123)
head(rt_preds)
glimpse(rt_preds)

# Step 2: For a single complete simulated dataset with same structure as original
rt_sim_data <- rt_preds %>%
  group_by(.row) %>%
  slice_sample(n = 1) %>%  # Take one random prediction per original observation
  ungroup() 
head(rt_sim_data)
glimpse(rt_sim_data)
```

## create summary rt data ##

at the pid level

```{r}
rt_sim_pid <- rt_sim_data %>% 
  group_by(PID, GROUP) %>%
  summarise(mean_rt = mean(.prediction, na.rm = TRUE), 
            sd_rt = sd(.prediction, na.rm = TRUE))
head(rt_sim_pid)
```

at the group level

```{r}
rt_sim_g <- rt_sim_data %>% 
  group_by(GROUP) %>%
  summarise(n=length(unique(PID)),
            mean_rt = mean(.prediction, na.rm = TRUE), 
            sd_rt = sd(.prediction, na.rm = TRUE), 
            sem_rt = (sd_rt/sqrt(n)))
rt_sim_g

# # A tibble: 4 × 5
#   GROUP         n mean_rt sd_rt sem_rt
#   <fct>     <int>   <dbl> <dbl>  <dbl>
# 1 untrained    18   1149.  380.   89.5
# 2 name         18   1072.  382.   90.0
# 3 tie          18   1123.  334.   78.7
# 4 both         18   1029.  301.   70.9
```

## plot rt ##

```{r}
p3.1 <- ggplot(rt_sim_pid, aes(x = GROUP, y = mean_rt, 
                               fill = GROUP, colour = GROUP)) +
  geom_jitter(position=position_jitterdodge(dodge.width =1), 
              size = 2.5, colour="grey") +
  geom_violin(alpha = 0.5, position=pd2) +
  geom_errorbar(data = rt_sim_g, aes(ymin=mean_rt-sem_rt,
                                         ymax=mean_rt+sem_rt), 
                width=.3, linewidth =0.5, colour = "black", position=pd2) +
  geom_point(data = rt_sim_g, colour = "black", position=pd2, size =4) +
  geom_line(data=rt_sim_g, group=1, colour = "black") +
  xlab("Training Condition") + 
  ylab('Mean Reaction Time (ms)') + 
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(800, 1600),
                     breaks = seq(800, 1600, 200))
p3.1
```

This also looks sensible, given the descriptives from the original data.

# 4. plot accuracy and rt together #

plot together

```{r}
p4.1 <- (p2.1 | p3.1) +
  plot_annotation(tag_levels = 'A')
p4.1

ggsave("figures/simulated/violin_acc_rt.jpeg",
       width = 10, height = 6, dpi = 800)
```

# 5. wrangle to create a simulated "raw" dataset #

here we combine simulated accuracy and reaction time data and wrangle it to have the same structure as the original data.

The purpose of doing so is so that folks can use this simulated data file to test our code.

## look at the actual raw data to get a sense of the structure ##

```{r}
raw <- read_csv("data/processed/day5_post_knots_data.csv")
head(raw)

# # A tibble: 6 × 14
#     PID RESPONSE    RT  STIM GROUP PICNOS1 PICNOS2 MATCH ORTN1 ORTN2 PICID1 PICID2   DAY SESSION
#   <dbl>    <dbl> <dbl> <dbl> <dbl>   <dbl>   <dbl> <dbl> <dbl> <dbl>  <dbl>  <dbl> <dbl>   <dbl>

colnames(raw)
# [1] "PID"      "RESPONSE" "RT"       "STIM"     "GROUP"    "PICNOS1"  "PICNOS2"  "MATCH"    "ORTN1"   
# [10] "ORTN2"    "PICID1"   "PICID2"   "DAY"      "SESSION" 
```

## wrangle the predicted accuracy data ##

```{r}
sim_raw_acc <- acc_sim_data |> 
  select(PID, GROUP, MATCH, .prediction) |> 
  mutate(RESPONSE = case_when(
    MATCH == 1 & .prediction == 1 ~ 1,  # MATCH 1, correct response
    MATCH == 1 & .prediction == 0 ~ 2,  # MATCH 1, incorrect response
    MATCH == 2 & .prediction == 1 ~ 2,  # MATCH 2, correct response
    MATCH == 2 & .prediction == 0 ~ 1,  # MATCH 2, incorrect response
    TRUE ~ NA_real_                    # Handle any other cases
  )) |> 
  select(-.prediction)
head(sim_raw_acc)
```

## wrangle the predicted reaction time data ##

```{r}
sim_raw_rt <- rt_sim_data |> 
  select(PID, GROUP, MATCH, .prediction) |> 
  rename(RT = .prediction)
head(sim_raw_rt)
```

## join them acc and rt together ##

```{r}
sim_raw_acc_rt <- sim_raw_acc |>
  mutate(RT = sim_raw_rt$RT)
head(sim_raw_acc_rt)
glimpse(sim_raw_acc_rt)
```

## join with the other raw data columns ##

```{r}
sim_raw <- raw |> 
  mutate(RESPONSE = sim_raw_acc$RESPONSE,
         RT = sim_raw_rt$RT)
head(sim_raw)
glimpse(sim_raw)
```

# 6. write out the simulated data #

```{r}
write_csv(sim_raw, "data/simulated/sim_raw.csv")
```